\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

\title{MIT 6.042J HW1 2010}
\author{Kelly Lin }
\date{June 2019}

\begin{document}

\maketitle

\section{Problem 1}
a. There are people who have taken 6.042 and have gotten A's in 6.042. 
For one person:
\begin {align*}
\exists x \in X. S(x) \wedge A(x)
\end {align*}
For at least two people:
\begin {align*}
\exists x, y \in X. S(x) \wedge A(x) \wedge S(y) \wedge A(y) \wedge \neg E(x,y)
\end {align*}
b. All people who are 6.042 TAs and have taken 6.042 got A's 6.042. 
\begin {align*}
\forall x \in X. T(x) \wedge \neg A(x) \rightarrow A(x)
\end {align*}
c. There are no people who are 6.042 TAs who did not get A's in 6.042.
\begin {align*}
\neg (\exists x \in X. (T(x) \wedge \neg A(x)))
\end {align*}
d. These are at least three people who are TAs in 6.042 and have not taken 6.042. 
\begin {align*}
\exists x,y,z \in X. (T(x) \wedge \neg S(x)) \wedge (T(y) \wedge \neg S(y)) \wedge (T(z) \wedge \neg S(z)) \wedge (\neg E(x,y) \wedge \neg E(y,z) \wedge \neg E(z,x))
\end {align*}
\section{Problem 2}
a. Use a truth table to prove or disprove the following:
\begin {align*}
\neg (P \vee (Q \wedge R)) = (\neg P) \wedge (\neg Q \vee \neg R)
\end {align*}
I made a truth table on scratch paper--not interested in reproducing it on here. 
This equation is true. There are 5 cases that are false and 3 that are true and they match up. 
\\\\
b.
\begin {align*}
\neg (P \wedge (Q \vee R)) = \neg P \vee (\neg Q \vee \neg R)
\end {align*}
This equation is false. The LHS has three false and 5 true. The RHS had 1 false and 7 true. 
\section{Problem 3}
Using nand, find an equivalent expression.
\\
i. $A \wedge B$
\\
$\neg (A \; nand \; B)$
\\\\
ii. $A \vee B$
\\
$(\neg A \; nand \; \neg B)$
\\
I actually came up with the nand version of it before I did the not version.
\\\\
iii. $A \rightarrow B$
\\
$A \; nand \; (\neg B)$
\\
For this one, we want something that will be true except when A is false and B is true. Because for all the other combos of A and B, they are valid. However when A is false and B is true, A implies B so if A is true and B is false that's not implying and the statement doesn't hold.
\\\\
b. Find an equivalent expression for $\neg A$ using only nands.
\\\\
$A \; nand \; A$ gives not A.
\\\\
c. Construct an expression using A and nand that evaluates to true regardless of whether A is true or false. Then do the same for one that evaluates to false.
\\\\
For true:
\\
$(A \; nand \; A) \; nand \; A$
\\\\
And for false:\\
$((A \; nand \; A) \; nand \; A) \; nand \; ((A \; nand \; A) \; nand \; A)$
\\
Basically just not-ing itself, since not is the self nand self construction as written above in part c. 
\\
\section{Problem 4}

\end{document}

